---
title: "MA585-HW6"
author: "Ranfei Xu"
date: "2022/3/23"
output: pdf_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(TSA)
library(xts)
library(tseries)
```

# 1

### a. Plot the data. Describe the features of the data. Do the data look stationary? Explain your answer.
```{r}
plot.ts(JohnsonJohnson, main='JohnsonJohnson')
```

This time series plot seems not stationary, since it has an obvious trend pattern and the variance seems to change over the time.

### b. Apply an appropriate variance stabilizing transformation, if necessary.
```{r}
x <- log(JohnsonJohnson)
plot.ts(x, main = 'log(JohnsonJohnson)')
```

After I apply a log transformation, the variance seems to be stabilized. 

### c. Carry out classical decomposition of the data, plot the transformed series along with the ACF and PACF.
```{r}
z <- decompose(x, type="additive")

# decomp.plot function from lecture notes
decomp.plot <-  function(x, main = NULL, ...) 
{ 
  if(is.null(main)) 
    main <-  paste("Decomposition of", x$type, "time series") 
  plot(cbind(observed = x$random + if (x$type == "additive") 
    x$trend + x$seasonal 
    else x$trend * x$seasonal, trend = x$trend, seasonal = x$seasonal, 
    random = x$random), main = main, ...) 
} 
decomp.plot(z,main="Additive Decomposition of log(JohnsonJohnson) Data") 

# plot ACF and PACF
par(mfrow=c(1,2))
acf(z$random, na.action = na.pass, lag.max=20, main = 'ACF of Random Part')
pacf(z$random, na.action = na.pass, lag.max=20, main = 'PACF of Random Part')
```


### d. Identify an ARMA model for the transformed data.



```{r}
# Check with EACF
TSA::eacf(na.omit(z$random))
```
According to the ACF and PACF plots, we can see ACF cuts-off after lag 1, and PACF cuts-off after lag 3. So the model might be  MA(1), AR(3), or ARMA(3,1). Based on the result of EACF table, the model is more likely to be MA(1).



### e. Repeat c and d, but instead of classical decomposition, use differencing to make the data stationary.

```{r}
# differencing with k=1
x1 <- diff(x)
plot.ts(x1, main = 'log(JohnsonJohnson) (Differencing k=1)')

# plot ACF and PACF
par(mfrow=c(1,2)) 
acf(x1, na.action = na.pass, lag.max=20, main = 'ACF after Differencing(k=1)')
pacf(x1, na.action = na.pass, lag.max=20, main = 'PACF after Differencing(k=1)')

# check EACF
TSA::eacf(x1)
```

According to the ACF and PACF plots, we can see ACF decays to zero, and PACF cuts-off after lag 3 (significantly. So the model might be AR(3). The result of EACF table is a valid proof.

# 2 

Consider the time series of the numbers of users connected to the Internet through a server every minute (R data WWWusage). Carry out a test for unit root. Apply necessary transformation and identify plausible ARMA models.

```{r}
x <- WWWusage
adf.test(x)
```

The result of Augmented Dickey-Fuller Test shows that p-value = 0.310, which fails to reject H0, thus the process is not stationary. So I apply a log transformation for WWWusage data.

```{r}
# log transformation
x1 <- log(WWWusage)
adf.test(x1)
# still fail to reject H0, try differencing


```

The result of Augmented Dickey-Fuller Test shows that p-value = 0.2622, which still fails to reject H0, thus I apply differencing transformation to stationary the data.

```{r}
# apply differencing with k=2
x2 <- diff(diff(WWWusage))
adf.test(x2)
```

After apply second order difference, the result of Augmented Dickey-Fuller Test finally shows a p-value which is small enough to reject H0. Thus the process is stationary now.

```{r}
par(mfrow=c(2,2)) 
plot.ts(x2)
acf(x2, lag.max=20, main = 'ACF after Differencing (k=2)')
pacf(x2, lag.max=20, main = 'ACF after Differencing (k=2)')
```

According to the ACF and PACF plots, the WWWusage data after second order difference might be an ARMA(0,0) process, which means it might be a white noise process.



# 6 (f.) 

Write a R code generate a random sample of size 500 from the AR(1) process with phi = 0.6 and sigma^2 = 0.8, Plot the simulated series along with the sample ACF and PACF of the series. Is the sample ACF and PACF consistent with AR(1)?
```{r}
# simulate AR(1) process
AR1.sm <- list(order = c(1,0,0), ar = 0.6, sd = sqrt(0.8))
x <- arima.sim(n=500, AR1.sm)
plot.ts(x, main = 'simulated AR(1)')
par(mfrow=c(1,2))
acf(x, main = 'ACF of simulated series')
pacf(x, main = 'PACF of simulated series')
```

As we can see, the ACF decay to zero with a damped sine wave, and the PACF cuts-off after lag 1, that exactly consistent with AR(1) process.
