---
title: "MA585-HW1"
author: "Ranfei Xu"
date: "2022/1/31"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.width=8, fig.height=3)
library(TSA)
```

# 1.

Based on Google Trends (https://www.google.com/trends/), think of three search terms each with the following characteristics:

```{r}
ted <- read.csv("E:/MA585/HW585/TED_multiTimeline.csv", header = TRUE, skip = 2)
ski <- read.csv("E:/MA585/HW585/ski_multiTimeline.csv", header = TRUE)
nfl <- read.csv("E:/MA585/HW585/netflix_multiTimeline.csv", header = TRUE, skip = 2)
```


(1) a time series with a clear trend component

The plot below shows the frequency of search term **TED**. We can see that there is a clear downward trend, which indicates that with the prosperity of video platforms, the popularity of ted has gradually declined.

```{r}
# head(ted)
plot.ts(ted$TED...United.States., ylab="Frequency",xlab="Week")
title("Daily Google Search for TED in the Past 5 Years")
```



(2) a time series with a clear seasonal component.

The plot below shows the frequency of search term **ski**. We can see that there is a clear seasonal component, which indicates Skiing is only widely participated in winter.

```{r}
# head(ski)
plot.ts(ski$ski...United.States., ylab="Frequency",xlab="Week")
title("Daily Google Search for ski in the Past 5 Years")
```

(3) a time series with no clear trend or seasonal pattern and is poorly described
by a combination of trend and seasonal components.

The plot below shows the frequency of search term **Netflix**. We can see that there is no clear trend or seasonal pattern. 

```{r}
# head(nfl)
plot.ts(nfl$Netflix...United.States., ylab="Frequency",xlab="Day")
title("Daily Google Search for netflix in the Past 90 Days")
```

\newpage

# 2. 


|Col. | Description |
| --- |:-------------:|
|TPCP | Total precipitation amount for the month (tens of mm) |
|TSNW | Total snowfall amount for the month (mm)|
|MMXT | Monthly mean maximum temperature (tenths of degrees of Celsius)|
|MMNT | Monthly mean temperature (tenths of degrees of Celsius)|

(1) Plot the time series for each variable and write a short description of the key features of each series. 

```{r}
BosClimateData <- read.csv("E:/MA585/HW585/BosClimateData.csv", 
header = TRUE)

BosData <- subset(BosClimateData,BosClimateData$STATION_NAME=="BOSTON LOGAN INTERNATIONAL AIRPORT MA US")

plot.ts(BosData$TPCP, xlab = "Month")
title("Total precipitation amount for the month (tens of mm)")
```

#### _Features_: 

The plot above describes the total precipitation amount for each month during 2020.01 to 2015.12. However there is no clear trend or seasonal pattern.


```{r}
plot.ts(BosData$TSNW, xlab = "Month")
title("Total snowfall amount for the month (mm)")
```

#### _Features_: 

The plot above describes the total snowfall amount for each month during 2020.01 to 2015.12. We can see a clear seasonal component that indicates Snowfall is only concentrated at certain times of the year.


```{r}
plot.ts(BosData$MMXT, xlab = "Month")
title("Monthly mean maximum temperature (tenths of degrees of Celsius)")
```

#### _Features_: 

The plot above describes the monthly mean maximum temperature during 2020.01 to 2015.12. We can see a clear seasonal component with continuity, which indicates the monthly mean maximum temperature in Boston changes stable, continuous and noticeable.

```{r}
plot.ts(BosData$MMNT, xlab = "Month")
title("Monthly mean temperature (tenths of degrees of Celsius)")
```

#### _Features_: 

The plot above describes the monthly mean temperature during 2020.01 to 2015.12. We can see a clear seasonal component with continuity, just as the Monthly mean maximum temperature plot.


(2) What proportion of months had total snowfall greater than a foot? 

```{r}
# 1 foot = 304.8 mm
TSNW   <- BosData$TSNW
l.TSNW <- length(TSNW)
s.TSNW <- length(subset(BosData$TSNW, BosData$TSNW >= 304.8))
(s.TSNW/l.TSNW)
```


(3) What proportion of months mean maximum temperature exceeded 80F?

```{r}
#  80F = 26.67 Celsius, 
MMXT   <- BosData$MMXT
l.MMXT <- length(MMXT)
s.MMXT <- length(subset(BosData$MMXT, BosData$MMXT >= 266.7))
(s.MMXT/l.MMXT)
```

\newpage

# 3. Simulation

(1) 
```{r}
plot.ts(rnorm(100))
```

#### _Answer_: 

After repeat the simulation several times, there are no obvious patterns in the plots, so I think it looks random.

(2)

```{r}
plot.ts(rt(n=100, df=4))
```

#### _Answer_: 

The amplitude is small, which means the	variance is relatively small at most of the "Time", thus might indicates the data generating process may not be Gaussian.